---
title: "Mini Data Analysis Milestone 2_TB"
author: "Tiana Broen, 16922171"
date: "19/10/2021"
output: github_document
---

```{r load-packages, include=FALSE}
library(datateachr)
library(tidyverse)
```
##### Introduction: This file contains all components of Milestone 2 of the Mini Data Analysis for STAT 545A. In this code, the data *apt_buildings* is further examined and cleaned to produce a reproducible and clear report using R Markdown. The dataset being used is from the *datateachr* package by Hayley Boyce and Jordan Bourak.

## Task 1: Process and summarize your data
### 1.1: Write the 4 research questions you defined in milestone 1.
The four research questions I have chosen to explore using the "apt_buildings" dataset are:
1. What property type (i.e. private, TCHC, Social Housing) is most likely to allow pets?
2. Is there a relationship between the accessibility of apartments and the year they were built?
3. What property type (i.e. private, TCHC, Social Housing) would be most appropriate for a family (i.e. including child play area, parking, guest parking, laundry room, and storage)*
4. Are apartments that allow pets also more likely to offer parking?

## 1.2: Choose one task from **Summarizing** and one task from **Graphing** to complete for each of the above four research questions.
### Research Question #1: "What property type is most likely to allow pets?"
**Task 4 from Summarizing:**: Based on two categorical variables, calculate two summary statistics of your choosing
For this task, I am interested in summary statistics for the "pets_allowed" and "property_type" categorical variables. First, I want to understand how many buildings allow/do not allow pets and the number of buildings belonging to each housing property category (private properties, social housing properties, or TCHC properties). Next, I count the number of buildings belonging to each property type that allows pets. This will allow me to calculate which property type is most likely to allow pets in my future analyses.  
```{r}
(pet_friendly_buildings_count <- apt_buildings %>% group_by(pets_allowed) %>%
  summarise(count=n())) #In total, 601 buildings do not allow pets, 2764 buildings allow pets, and 90 do not specify whether or not they allow pets.

(property_type_count <- apt_buildings %>% group_by(property_type) %>%
  summarise(count=n())) #In total, 2888 buildings are private properties, 240 buildings are social housing properties, and 327 are TCHC properties (Toronto Community Housing Corporation)

(pet_friendly_only <- apt_buildings %>% 
    group_by(property_type) %>%
    filter(pets_allowed == "YES") %>%
    summarise(count=n()))#This tells me how many buildings allow pets based on the property type. Based on this summary, the majority of pet friendly buildings are private properties.
```

**Task 5 from Graphing:** Create a graph out of summarized variables that has at least two geom layers.
I selected a bar graph to represent the summarized data about "property_type" and "pets_allowed". The bar graph allows me to easily visualize the number of pet friendly buildings in each property type as well as the general ratio of Pets Allowed vs. Not Allowed vs. NA within each property type.
```{r}
ggplot_pet_friendly <- ggplot(data = apt_buildings) + 
  geom_bar(mapping = aes(x = property_type, fill = pets_allowed), position = "dodge") +
  xlab("Property Type") + ylab("Number of Buildings")
(ggplot_pet_friendly + labs(fill="Pets Allowed?"))
```

### Research Question #2: "Is there a relationship between the accessibility of apartments and the year they were built?"
**Task 1 from Summarizing:**: Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.
I am interested in better understanding the accessibility of buildings and the years they were built. First, I create a subset of data including buildings that have "barrier free accessibility entrances". Next, within this subset, I calculate four summary statistics for the years these buildings were built (range, mean, median, sd). This will help me get a general idea of the years that buildings with accessible entrances were built.
```{r}
accessible <- apt_buildings %>% 
  filter(barrier_free_accessibilty_entr == "YES")

(summary_statistics_year_built <- summary(accessible$year_built))
(sd_year_built <- sd(accessible$year_built, na.rm=TRUE))
```

**Task 7 from Graphing:**: Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are “pretty” or easier to read.
I chose to represent the relationship between number of accessible units and year built using a scatter plot graph. I made the y-axis logarithmic in order to view the heavily skewed data more clearly. This allowed me to visualize that most buildings with a large number of accessible units were built after 1950. However, since there are several buildings with zero accessible units, the graph visualization is not very clear. In the future, I could filter my data to exclude buildings that do not have any accessible units.
```{r}
(ggplot_accessible_year_built <- ggplot(data = apt_buildings) + 
  geom_point(mapping = aes(x = year_built, y = no_barrier_free_accessible_units)) +
  scale_y_continuous(trans = 'log10') + xlab("Year Built") + ylab("Number of Barrier Free Accessible Units"))
```

### Research Question #3: "What property type (i.e. private, TCHC, Social Housing) would be most appropriate for a family (i.e. including child play area, parking, guest parking, laundry room, and storage)?"
**Task 2 from Summarizing:** Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.
Since I am interested in properties that would be appropriate for families, it is important to understand the amenities, such as parking stalls, available to tenants. I am interested in the number of accessible parking spaces as it will tell me whether these properties will facilitate parents that drive children to school, carpool, or invite other friends over for play dates. As a starting point, I am interested in the average number of parking stalls available across all apartment buildings. In future analyses, I can subset me data based on other criteria a family may have for a home (ex. laundry room, storage). In this activity, I calculate range, mean, median, and standard deviation. 
```{r}
(summary_statistics_no_parking_spaces <- summary(apt_buildings$no_of_accessible_parking_spaces))
(sd_no_parking_spaces <- sd(apt_buildings$no_of_accessible_parking_spaces, na.rm=TRUE))
```

**Task 8 from Graphing:** Create 3 histograms out of summarized variables, with each histogram having different sized bins. Pick the "best" one and explain why it is the best.
I chose a histogram to visualize the number of parking stalls available across all apartment buildings. Since the number of parking stalls ranges from 0-340, I chose to include bin numbers of the maximum divided by 4 (85), ten (34), and 20 (17). Since the data is very right skewed, including too many bins makes it difficult to visualize the number of parking stalls available at smaller increments. I think the lowest bin number (17) is the best for visualizing the data as it allows you to view the general trend of the distribution of number of parking stalls, but still understand some finer details within the general trend (i.e. difference between number of apartments with 50 vs. 100 parking stalls). In future analyses and visulaization I could remove outliers from the data to create a more normal distribution and better reflect the mean of the data on number of accessible parking stalls.
```{r}
ggplot(apt_buildings, aes(no_of_accessible_parking_spaces)) +
  geom_histogram(bins = 85)

ggplot(apt_buildings, aes(no_of_accessible_parking_spaces)) +
  geom_histogram(bins = 34)

ggplot(apt_buildings, aes(no_of_accessible_parking_spaces)) +
  geom_histogram(bins = 17) 
```

### Research Question #4: "Are apartments that allow pets also more likely to offer parking?"
**Task 4:** Based on two categorical variables, calculate two summary statistics of your choosing.
As a dog owner, I was curious if apartments that allow pets also allow parking for when you need to travel with your pet. To explore these two categorical variables, I first wanted to understand the types of parking available across all apartments. I realized that there are several different combinations of parking types that are available at apartments. In the future, I could create new variables for each parking type to count the number of apartments with each type of parking. For these summary statistics, I chose to create a dichotomous variable of either "parking available" or "parking not available" by filtering out apartments with "NA" under *parking type*. Next, I wanted to better understand the relationship between each parking type and the pet-friendly status of the buildings by counting the number of buildings with parking available and pets-allowed.
```{r}
apt_buildings %>% group_by(parking_type) %>%
  summarise(count=n())
parking_available <- apt_buildings %>% filter(!is.na(parking_type))

(parking_and_pets <- parking_available %>% 
    filter(pets_allowed == "YES") %>%
    summarise(count=n())) #In total, 2555 buildings are both pet-friendly and offer a form of parking.
apt_buildings %>% summarise(count=n()) #There are a total of 3455 apartment buildings
(parking_and_pets_ratio <- 2555/3455) #73% of all apartments are both pet-friendly and offer parking.
```

**Task 7:** Create a graph out of summarized variables that has at least two geom layers.
Since my fourth research questions was fairly straightforward, I chose to explore my interest in pet-friendly buildings more. My research question here is: "What is the relationship between the pet-friendly status of apartments and number of storeys in apartment buildings?" I chose a bar graph to visualize this relationship, however I did not find this visualization particularly useful. In the future, I may choose to further explore whether buildings with fewer storeys are more likely to allow pets.
```{r}
(ggplot_pets_and_storeys <- ggplot(data = apt_buildings) + 
  geom_bar(mapping = aes(x = no_of_storeys, fill = pets_allowed)) +
  xlab("Number of Storeys") + ylab("Number of Buildings"))
(ggplot_pets_and_storeys + labs(fill="Pets Allowed?"))
```

### 1.3
Based on the above operations, I have definitely become closer to answering my research questions. Specifically, I was able to better visualize answers to my questions using ggplot2. Above each activity I have explained my reasoning for each analysis, the challenges I encountered, and my suggestions for future directions in analyzing and visualizing my data. 

# Task 2: Tidy your data

### 2.1
Since my dataset has over 8 variables, I have chose the following 8 variables to examine for tidy vs. untidy data:
1. fire_alarm
2. no_of_elevators
3. heating_type
4. pets_allowed
5. property_type
6. year_registered
7. year_built
8. no_of_units
Based on these 8 variables, my data is tidy. However, some variables not listed above do not meet the definition of tidy data. For example, the column *parking_type* is categorical and some rows include several values separated by a ",". In order to clean columns like this, I must split the column into multiple columns so that each column is only one variable. This will also allow me to create a consistent data structure so that I can better use my data analyzing tools in the future.

### 2.2
I will untidy the data and re-tidy it, showing a clear before and after.
```{r}
tidy_tibble <- apt_buildings %>%
  select(c(fire_alarm, no_of_elevators, heating_type, pets_allowed, property_type,
           year_registered, year_built, no_of_units)) #creating a tibble with the 8 variables I will untidy

#Before untidying:
tidy_apt_buildings <- tidy_tibble

#Untidying.I will pivot wider the *property type* variable, with the values indicating the type of heating. 
(untidy_heating_type <- tidy_apt_buildings %>%
    pivot_wider(names_from = property_type, values_from = heating_type, values_fill = NA))
untidy_heating_type %>%
  unnest

#Re-tidying
(tidy_heating_type <- untidy_heating_type %>%
    pivot_longer(cols = c('PRIVATE', 'TCHC', 'SOCIAL HOUSING'), names_to = "property_type",
                 values_to = "heating_type"))
tidy_heating_type %>%
  unnest
```

### 2.3
The two questions I am choosing to move forward with are:
1. *Is there a relationship between the accessibility of apartments and the year they were built?* I chose this question because I was not able to fully it in the above analyses. After exploring the data for this question I also understand that I can explore more about the accessibility of apartment buildings, such as accessible parking spaces and number of storeys in the building.
2. *What property type (i.e. private, TCHC, Social Housing) would be most appropriate for a family (i.e. including child play area, parking, guest parking, laundry room, and storage)* I am choosing to further explore this question because there are several more relationships I can examine to determine the most appropriate property type for a family. There are several requirements that families may look for in housing, and I can explore these requirements both separately and in relation to one another.

<<<<<<< HEAD:Milestone2_TB.Rmd
I will choose an appropriate version of my data to answer these two questions below
```{r}
(apt_buildings_final <- apt_buildings[c(1:3, 5, 6, 12:16, 18, 25, 30, 34, 37)]) #dropping irrelevant columns
(apt_buildings_final %>% filter(no_barrier_free_accessible_units > 0, na.rm=TRUE)) #Filtering out any buildings that do not offer barrier free accessible units as this is important for both of my above research questions.
(apt_buildings_final %>% rename(
  accessible_units = no_barrier_free_accessible_units,
  storage_available = locker_or_storage_room,
  accessible_entrance = barrier_free_accessibilty_entr)) #Re-naming columns so that they make more sense for future analyses
(apt_buildings_final %>% drop_na()) #Dropping all rows containing missing values as they will make it difficult to fully answer my research questions.
```
=======
### 2.3 (5 points)

Now, you should be more familiar with your data, and also have made progress in answering your research questions. Based on your interest, and your analyses, pick 2 of the 4 research questions to continue your analysis in milestone 3, and explain your decision. 

Try to choose a version of your data that you think will be appropriate to answer these 2 questions in milestone 3. Use between 4 and 8 functions that we've covered so far (i.e. by filtering, cleaning, tidy'ing, dropping irrelvant columns, etc.). 

<!--------------------------- Start your work below --------------------------->


<!----------------------------------------------------------------------------->

*When you are done, knit an `md` file. This is what we will mark! Make sure to open it and check that everything has knitted correctly before submitting your tagged release.*
>>>>>>> 140efce732c40f5438025095f20b88ba11c1c18c:Milestone_2/Milestone2_TB.Rmd
