---
title: "Milestone_3_TB"
author: "Tiana Broen, 16922171"
date: "27/10/2021"
output: github_document
---

##### Introduction: This file contains all components of Milestone 3 of the Mini Data Analysis for STAT 545A. In this code, the data *apt_buildings* is further examined and cleaned to produce a reproducible and clear report using R Markdown. The dataset being used is from the *datateachr* package by Hayley Boyce and Jordan Bourak.

The two research questions I will be examining from Milestone 2 are:  
1. Is there a relationship between the accessibility of apartments and the year they were built?  
2. What property type (i.e. private, TCHC, Social Housing) would be most appropriate for a family (i.e. including child play area, parking, guest parking, laundry room, and storage)

## Exercise 1: Special Data Types

I will use the following plot from Milestone 2 to complete tasks 1 & 2. The plot below examines the relationship between the year an apartment building was built and the number of accessible units in that building. In order to make this graph more applicable for this exercise, I have grouped the "year built" variable into "century built" and transformed it into a factor.
```{r}
apt_buildings_1 <- apt_buildings[c('year_built', 'no_barrier_free_accessible_units')] #new dataframe with only the variables I need for this exercise
apt_buildings_1$century_built <- apt_buildings_1$year_built
apt_buildings_1 <- filter(apt_buildings_1, !is.na(year_built))
apt_buildings_1 <- filter(apt_buildings_1, !is.na(no_barrier_free_accessible_units))

# transforming year built into century built
apt_buildings_1$century_built[c(apt_buildings_1$century_built > 1800 & apt_buildings_1$century_built <= 1850)] <- "1800-1850"
apt_buildings_1$century_built[c(apt_buildings_1$century_built > 1850 & apt_buildings_1$century_built <= 1900)] <- "1851-1900"
apt_buildings_1$century_built[apt_buildings_1$century_built > 1900 & apt_buildings_1$century_built <= 1950] <- "1901-1950"
apt_buildings_1$century_built[c(apt_buildings_1$century_built > 1950 & apt_buildings_1$century_built <= 2000)] <- "1951-2000"
apt_buildings_1$century_built[apt_buildings_1$century_built > 2000] <- "2000+"
apt_buildings_1$century_built <- as.factor(apt_buildings_1$century_built)

#building box plot for tasks 1&2 below
(ggplot_1 <- ggplot(apt_buildings_1, aes(x=century_built, y=no_barrier_free_accessible_units, group=century_built)) +
  geom_boxplot(aes(fill=century_built)) + scale_y_continuous(trans='log10') +
    xlab("Century Built") + ylab("Number of Barrier Free Accessible Units") + labs(fill="Century Built"))
```

### Task 1: Produce a new plot that reorders a factor in your original plot.
```{r}
(ggplot_2 <- apt_buildings_1 %>%
  mutate(century_built = fct_infreq(century_built)) %>%
  ggplot(aes(x=century_built, y=no_barrier_free_accessible_units, group=century_built)) +
  geom_boxplot(aes(fill=century_built)) + scale_y_continuous(trans='log10') +
    xlab("Century Built") + ylab("Number of Barrier Free Accessible Units") + labs(fill="Century Built"))
```
I chose to reorder the factor "century built". I chose to reorder this in by infrequency, meaning I ordered this factor by the frequency in which it appears (i.e. most apartment buildings are built between 1951-2000, the fewest were built in the 1800-1850). This adds another layer of understanding to this plot as you can tell by a glance that although most apartment buildings were built in the 90s, more buildings were built in the 2000s with accessibility in mind.

### Task 2: Produce a new plot that groups some factor levels together into an "other" category.
```{r}
(ggplot_3 <- apt_buildings_1 %>%
   mutate(century_built = fct_infreq(century_built)) %>%
  mutate(century_built = fct_lump(century_built, n=1, w=NULL, other_level = "Other", ties.method = "min")) %>%
  ggplot(aes(x=century_built, y=no_barrier_free_accessible_units, group=century_built)) +
  geom_boxplot(aes(fill=century_built)) + scale_y_continuous(trans='log10') +
    xlab("Century Built") + ylab("Number of Barrier Free Accessible Units") + labs(fill="Century Built"))
```
 I chose to lump together the factors that showed up the least frequently in my data. This graph shows me that the majority of the buildings were built between 1951 and 2000. However, the average number of accessible units is slightly greater when you combine all other years into one factor.
 
## Exercise 2: Modelling

### 2.0: Pick a research question, and pick a variable of interest that's relevant to the research question.
**Research question**: Is there a relationship between the accessibility of apartments and the year they were built? 
**Variable of interest**: My variable of interest is: "number of barrier free accessible units".

### 2.1: Fit a model or run a hypothesis test that provides insight on this variable with respect to the research question. 
```{r}
model_1 <- lm(formula = no_barrier_free_accessible_units ~ year_built, data = apt_buildings)
model_1
```
### 2.2: Produce something relevant from your fitted model.

```{r}
(glance(model_1))
```
I chose to use the *broom* function *glance*. This gives me several pieces of information related to the model fit. Specifically, I was interested in the *r.squared* column. The *r.squared* column tells me how fitted the data is to the regression line; in this case, my data is not very closely fitted (0.009). So, this indicates that the relationship between number of barrier free accessible units and year built is not represented well by my above model. In the future, I should create a better fitting model (i.e. exclude outliers, look at correlation coefficients to understand my data better) to get more meaningful results.

## Exercise 3: Reading and writing data
```{r}
dir.create(here::here("C:\\Users\\tiana\\OneDrive\\Desktop\\output"))
```

### 3.1 (5 points)
```{r}
# Reformatting my summary table from Milestone 2, exercise 1.2
apt_buildings_na <- filter(apt_buildings, !is.na(year_built))
apt_buildings_na <- filter(apt_buildings_na, !is.na(no_barrier_free_accessible_units))
apt_buildings_na

sumstat <- apt_buildings_na %>%
    select(
        `Year Built` = year_built,
        `Number of Accessible Units` = no_barrier_free_accessible_units
        ) %>%
  summarise_each(funs(mean, sd, min, max)) %>%
    gather(key, value, everything()) %>% 
    separate(key, into = c("variable", "stat"), sep = "_") %>%
    spread(stat, value) %>%
    select(variable, mean, sd, min, max) %>%
    mutate_each(funs(round(., 1)), -variable)
sumstat

sumstat_df <- as.data.frame(describe(sumstat)) #formatting sumstat_df as a dataframe to export to .csv file
sumstat_df

write_csv(sumstat_df, here::here("output", "sumstat_df_exported_file.csv"))
dir(here::here("output"))
```

### 3.2: Write your model object from Exercise 2 to an R binary file (an RDS), and load it again.
```{r}
saveRDS(model_1, here::here("output", "model_1.rds"))
readRDS(here::here("output", "model_1.rds"))
dir(here::here("output"))
```
## File and Folder structure (3 points)

You should have at least four folders in the top level of your repository: one for each milestone, and one output folder. If there are any other folders, these are explained in the main README.

Each milestone document is contained in its respective folder, and nowhere else. 

Every level-1 folder (that is, the ones stored in the top level, like "Milestone1" and "output") has a `README` file, explaining in a sentence or two what is in the folder, in plain language (it's enough to say something like "This folder contains the source for Milestone 1").

## Output (2 points)

All output is recent and relevant: 

- All Rmd files have been `knit`ted to their output, and all data files saved from Exercise 3 above appear in the `output` folder. 
- All of these output files are up-to-date -- that is, they haven't fallen behind after the source (Rmd) files have been updated.
- There should be no relic output files. For example, if you were knitting an Rmd to html, but then changed the output to be only a markdown file, then the html file is a relic and should be deleted. 

Our recommendation: delete all output files, and re-knit each milestone's Rmd file, so that everything is up to date and relevant. 

PS: there's a way where you can run all project code using a single command, instead of clicking "knit" three times. More on this in STAT 545B!