---
title: "Milestone_3_TB"
author: "Tiana Broen, 16922171"
date: "27/10/2021"
output: github_document
---

```{r, message = FALSE}
library(datateachr) #specific data I am analyzing: "apt_buildings"
library(tidyverse)
```
##### Introduction: This file contains all components of Milestone 3 of the Mini Data Analysis for STAT 545A. In this code, the data *apt_buildings* is further examined and cleaned to produce a reproducible and clear report using R Markdown. The dataset being used is from the *datateachr* package by Hayley Boyce and Jordan Bourak.

The two research questions I will be examining from Milestone 2 are:  
1. Is there a relationship between the accessibility of apartments and the year they were built?  
2. What property type (i.e. private, TCHC, Social Housing) would be most appropriate for a family (i.e. including child play area, parking, guest parking, laundry room, and storage)

## Exercise 1: Special Data Types

I will use the following plot from Milestone 2 to complete tasks 1 & 2. The plot below examines the relationship between the year an apartment building was built and the number of accessible units in that building. In order to make this graph more applicable for this exercise, I have grouped the "year built" variable into "century built" and transformed it into a factor.
```{r}
apt_buildings_1 <- apt_buildings[c('year_built', 'no_barrier_free_accessible_units')] #new dataframe with only the variables I need for this exercise
apt_buildings_1$century_built <- apt_buildings_1$year_built
apt_buildings_1 <- filter(apt_buildings_1, !is.na(year_built))
apt_buildings_1 <- filter(apt_buildings_1, !is.na(no_barrier_free_accessible_units))

# transforming year built into century built
apt_buildings_1$century_built[c(apt_buildings_1$century_built > 1800 & apt_buildings_1$century_built <= 1850)] <- "1800-1850"
apt_buildings_1$century_built[c(apt_buildings_1$century_built > 1850 & apt_buildings_1$century_built <= 1900)] <- "1851-1900"
apt_buildings_1$century_built[apt_buildings_1$century_built > 1900 & apt_buildings_1$century_built <= 1950] <- "1901-1950"
apt_buildings_1$century_built[c(apt_buildings_1$century_built > 1950 & apt_buildings_1$century_built <= 2000)] <- "1951-2000"
apt_buildings_1$century_built[apt_buildings_1$century_built > 2000] <- "2000+"
apt_buildings_1$century_built <- as.factor(apt_buildings_1$century_built)

#building box plot for tasks 1&2 below
(ggplot_1 <- ggplot(apt_buildings_1, aes(x=century_built, y=no_barrier_free_accessible_units, group=century_built)) +
  geom_boxplot(aes(fill=century_built)) + scale_y_continuous(trans='log10') +
    xlab("Century Built") + ylab("Number of Barrier Free Accessible Units") + labs(fill="Century Built"))
```

### Task 1: Produce a new plot that reorders a factor in your original plot.
```{r}
(ggplot_2 <- apt_buildings_1 %>%
  mutate(century_built = fct_infreq(century_built)) %>%
  ggplot(aes(x=century_built, y=no_barrier_free_accessible_units, group=century_built)) +
  geom_boxplot(aes(fill=century_built)) + scale_y_continuous(trans='log10') +
    xlab("Century Built") + ylab("Number of Barrier Free Accessible Units") + labs(fill="Century Built"))
```
I chose to reorder the factor "century built". I chose to reorder this in by infrequency, meaning I ordered this factor by the frequency in which it appears (i.e. most apartment buildings are built between 1951-2000, the fewest were built in the 1800-1850). This adds another layer of understanding to this plot as you can tell by a glance that although most apartment buildings were built in the 90s, more buildings were built in the 2000s with accessibility in mind.

### Task 2: Produce a new plot that groups some factor levels together into an "other" category.
```{r}
(ggplot_3 <- apt_buildings_1 %>%
   mutate(century_built = fct_infreq(century_built)) %>%
  mutate(century_built = fct_lump(century_built, n=1, w=NULL, other_level = "Other", ties.method = "min")) %>%
  ggplot(aes(x=century_built, y=no_barrier_free_accessible_units, group=century_built)) +
  geom_boxplot(aes(fill=century_built)) + scale_y_continuous(trans='log10') +
    xlab("Century Built") + ylab("Number of Barrier Free Accessible Units") + labs(fill="Century Built"))
```
 I chose to lump together the factors that showed up the least frequently in my data. This graph shows me that the majority of the buildings were built between 1951 and 2000. However, the average number of accessible units is slightly greater when you combine all other years into one factor.
 
## Exercise 2: Modelling

### 2.0 (no points)

Pick a research question, and pick a variable of interest (we'll call it "Y") that's relevant to the research question. Indicate these.

<!-------------------------- Start your work below ---------------------------->

**Research Question**: FILL_THIS_IN

**Variable of interest**: FILL_THIS_IN

<!----------------------------------------------------------------------------->

## 2.1 (5 points)

Fit a model or run a hypothesis test that provides insight on this variable with respect to the research question. Store the model object as a variable, and print its output to screen. We'll omit having to justify your choice, because we don't expect you to know about model specifics in STAT 545. 

- **Note**: It's OK if you don't know how these models/tests work. Here are some examples of things you can do here, but the sky's the limit. 
    - You could fit a model that makes predictions on Y using another variable, by using the `lm()` function.
    - You could test whether the mean of Y equals 0 using `t.test()`, or maybe the mean across two groups are different using `t.test()`, or maybe the mean across multiple groups are different using `anova()` (you may have to pivot your data for the latter two).
    - You could use `lm()` to test for significance of regression.

<!-------------------------- Start your work below ---------------------------->


<!----------------------------------------------------------------------------->

## 2.2 (5 points)

Produce something relevant from your fitted model: either predictions on Y, or a single value like a regression coefficient or a p-value. 

- Be sure to indicate in writing what you chose to produce. 
- Your code should either output a tibble (in which case you should indicate the column that contains the thing you're looking for), or the thing you're looking for itself.
- Obtain your results using the `broom` package if possible. If your model is not compatible with the broom function you're needing, then you can obtain your results by some other means, but first indicate which broom function is not compatible. 
  
<!-------------------------- Start your work below ---------------------------->


<!----------------------------------------------------------------------------->



# Exercise 3: Reading and writing data

Get set up for this exercise by making a folder called `output` in the top level of your project folder / repository. You'll be saving things there. 

## 3.1 (5 points)

Take a summary table that you made from Milestone 2 (Exercise 1.2), and write it as a csv file in your `output` folder. Use the `here::here()` function.

- **Robustness criteria**: You should be able to move your Mini Project repository / project folder to some other location on your computer, or move this very Rmd file to another location within your project repository / folder, and your code should still work.
- **Reproducibility criteria**: You should be able to delete the csv file, and remake it simply by knitting this Rmd file. 
  
<!-------------------------- Start your work below ---------------------------->


<!----------------------------------------------------------------------------->

## 3.2 (5 points)

Write your model object from Exercise 2 to an R binary file (an RDS), and load it again. Be sure to save the binary file in your `output` folder. Use the functions `saveRDS()` and `readRDS()`. 

- The same robustness and reproducibility criteria as in 3.1 apply here.

<!-------------------------- Start your work below ---------------------------->


<!----------------------------------------------------------------------------->

# Tidy Repository

Now that this is your last milestone, your entire project repository should be organized. Here are the criteria we're looking for.

## Main README (3 points)

There should be a file named `README.md` at the top level of your repository. Its contents should automatically appear when you visit the repository on GitHub. 

Minimum contents of the README file:

- In a sentence or two, explains what this repository is, so that future-you or someone else stumbling on your repository can be oriented to the repository.
- In a sentence or two (or more??), briefly explains how to engage with the repository. You can assume the person reading knows the material from STAT 545A. Basically, if a visitor to your repository wants to explore your project, what should they know?

Once you get in the habit of making README files, and seeing more README files in other projects, you'll wonder how you ever got by without them! They are tremendously helpful.

## File and Folder structure (3 points)

You should have at least four folders in the top level of your repository: one for each milestone, and one output folder. If there are any other folders, these are explained in the main README.

Each milestone document is contained in its respective folder, and nowhere else. 

Every level-1 folder (that is, the ones stored in the top level, like "Milestone1" and "output") has a `README` file, explaining in a sentence or two what is in the folder, in plain language (it's enough to say something like "This folder contains the source for Milestone 1").

## Output (2 points)

All output is recent and relevant: 

- All Rmd files have been `knit`ted to their output, and all data files saved from Exercise 3 above appear in the `output` folder. 
- All of these output files are up-to-date -- that is, they haven't fallen behind after the source (Rmd) files have been updated.
- There should be no relic output files. For example, if you were knitting an Rmd to html, but then changed the output to be only a markdown file, then the html file is a relic and should be deleted. 

Our recommendation: delete all output files, and re-knit each milestone's Rmd file, so that everything is up to date and relevant. 

PS: there's a way where you can run all project code using a single command, instead of clicking "knit" three times. More on this in STAT 545B!

## Error-free code (1 point)

This Milestone 3 document knits error-free. (We've already graded this aspect for Milestone 1 and 2)

## Tagged release (1 point)

You've tagged a release for Milestone 3. (We've already graded this aspect for Milestone 1 and 2)